\pkg{SVMBridge}: Using SVM Solvers from R

### 2015-09-16 paper/article.tex
snapshot

% TODO:
% C++/R -> prolang{.}
% SVMBridge -> pkg{.}

@@

%% an abstract and keywords
\Abstract{
  Most SVM Solver do not come with a wrapper in R.
  In case the SVM Solver is written in C++, it can be linked to
  a package via Rcpp. Although possible, this entails a lot of work to do,
  and is not possible, if the SVM Solver is written in other languages.
  Alternatively, one can call the SVM Solver from within R by a system command.
  SVMBridge eases this calls by providing a framework and ready wrappers
  for several SVM Solvers like LASVM, SVMperf, LLSVM and BVM/CVM.  
}
\Keywords{support vector machines, command line}
\Plainkeywords{support vector machines, command line} %% without formatting
%% at least one keyword must be supplied

@@

\subsection{SVM packages in R}

There are only a few SVM libraries readily available from within R.
This includes the kernlab, the e1071 and the lasvmR packages.
All of these link existing C++ sources directly from R and need
more or less extensive wrappers to allow the user direct access to
the options of the underlying SVM solver.

Other SVM solver, like the BudgetedSVM package~\citep{djuric2014budgetedsvm}, 
which contains BSGD and LLSVM, or SVMperf~\citep{joachims2009sparse}, cannot be called directly from R.
Instead, a system command has to be issued.
To ease this calls, the \pkg{SVMBridge} package has been developed.

@@

@@

A wrapper consists of several routines: It must provide routines
for assembling the command line call for training as well as testing,
reading and writing the model and a few general routines like searching
the binaries and printing.

@@

Although reading these files into R is possible via either
the \pkg{kernlab} or the \pkg{e1071} routines, both only provide R solutions,
and therefore suffer from suboptimal performance.
The SVMBridge package provides a simple sparse data format 
reading and writing, which are implemented in C++ and
therefore are nearly two orders of magnitudes faster than
the corresponding e1071 routines.
Notice, that currently only dense matrices are supported,
TODO: work around this with the Sparse Matrix Package.

@@

Multiclass is supported, where possible.
Note that several SVM packages only support binary problems.
In these cases, training a one-vs-all machine is rather easy, if the
data is loaded into R first.

@@

Every wrapper should also be able to find tiself in a given directory tree.
For this \code{findSoftware} is provided.

At last, for querying reasons, there is also a print routine
so that wrapper can print out any further information about itself.

@@

We provided a simple framework to attach many SVM solvers to the R Subsystem.
This allows for a systematic call of them.
By providing a wrapper, linking any SVM solver to R become much easier.

@@

We acknowledge support by the Mercator Research Center Ruhr,
under grant Pr-2013-0015 \textit{Support-Vektor-Maschinen f{\"u}r extrem gro{\ss}e Datenmengen} and
partial support by the German Research Foundation (DFG) within the Collaborative Research Centers SFB 823
\textit{Statistical modelling of nonlinear dynamic processes}, Project C2.

>>>>

% TODO:
% \proglang{C++}/R -> prolang{.}
% SVMBridge -> pkg{.}

@@

%% an abstract and keywords
\Abstract{
  Most SVM Solver do not come with a wrapper in  \proglang{R}.
  In case the SVM Solver is written in \proglang{C++}, it can be linked to
  a package via  Rcpp. Although possible, this entails a lot of work to do,
  and is not possible, if the SVM Solver is written in other languages.
  Alternatively, one can call the SVM Solver from within \proglang{R} by a system command.
  SVMBridge eases this calls by providing a framework and ready wrappers
  for several SVM Solvers like LASVM, SVMperf, LLSVM and BVM/CVM.  
}
\Keywords{support vector machines, command line}
\Plainkeywords{support vector machines, command line} %% without formatting
%% at least one keyword must be supplied

@@

\subsection[SVM packages in R]{SVM packages in \proglang{R}}

There are only a few SVM libraries readily available from within \proglang{R}.
This includes the well-known packages like \pkg{kernlab} (via \code{ksvm()})~\citep{kernlab}, 
and \pkg{e1071} (via \code{svm()})~\citep{dimitriadou2008misc}
as well as other less known ones, e.g. \pkg{SwarmSVM}~\citep{SwarmSVM} 
and \pkg{lasvmR}~\citep{lasvmR}.
All of these link existing \proglang{C++} sources directly from within \proglang{R} and need
more or less extensive wrappers to allow the user direct access to
the options of the underlying SVM solver.

Other SVM solvers, like the BudgetedSVM package~\citep{djuric2014budgetedsvm}, 
which contains BSGD and LLSVM, or SVMperf~\citep{joachims2009sparse}, cannot be called directly from  \proglang{R}.
Instead, a system command has to be issued.
To ease this calls, the \pkg{SVMBridge} package has been developed.

@@

The \pkg{SVMBridge} provides a framework to incorporate 
SVM solvers by means of direct calls (via \code{system2()} commands).
This is different to \pkg{e1071} and \pkg{kernlab}, which link (and
thus copy) existing \proglang{C++} code directly inside the package.
\pkg{SVMBridge} works thus differently: Each SVM solver must provide
its own socalled {\it Wrapper}. This is a piece of R code that contains
code e.g. to allow the proper calling of the system binary of a SVM solver
as well as handling the corresponding models. Such wrapper can be
registered in the SVMBridge and henceforth can be used opaquely.

@@

A wrapper is a S3 object: It must provide routines
for assembling the command line call for training as well as testing,
reading and writing the model and a few general routines like searching
the binaries and printing.

@@

Although reading these files into \proglang{R} is possible via either
the \pkg{kernlab} or the \pkg{e1071} routines, both only provide \proglang{R} solutions,
and therefore suffer from suboptimal performance.
The SVMBridge package provides a simple sparse data format 
reading and writing, which are implemented in \proglang{C++} and
therefore are nearly two orders of magnitudes faster than
the corresponding e1071 routines.
Notice, that currently only dense matrices are supported,
TODO: work around this with the Sparse Matrix Package.

@@

Multiclass is supported, where possible.
Note that several SVM packages only support binary problems.
In these cases, training a one-vs-all machine is rather easy, if the
data is loaded into \proglang{R} first.

@@

Every wrapper should also be able to find itself in a given directory tree.
For this the S3 method \code{findSoftware()} must be provided.

At last, for querying reasons, there is also a print routine
so that wrapper can print out any further information about itself.
To see the currently loaded wrapper, one can call 
\code{overviewAllSVMSolver()}.

@@

We provided a simple framework, \pkg{SVMBridge},  to attach many SVM solvers to the \proglang{R} landscape.
\pkg{SVMBridge} does this by an easily extensible wrappers that manage the exchange of data and models
via command line.
By providing such a wrapper, linking any SVM solver to \proglang{R} becomes much easier.

@@

We acknowledge support by the Mercator  \proglang{R}.search Center  \proglang{R}.hr,
under grant Pr-2013-0015 \textit{Support-Vektor-Maschinen f{\"u}r extrem gro{\ss}e Datenmengen} and
partial support by the German  \proglang{R}.search Foundation (DFG) within the Collaborative  \proglang{R}.search Centers SFB 823
\textit{Statistical modelling of nonlinear dynamic processes}, Project C2.
